---
title: "Lactate Testing Report"
author: "Jeff Shilt"
date: "2023-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=FALSE, warning=FALSE, echo=FALSE}

#Load Libraries and Data
library(tidyverse)
library(here)
library(lactater)

Data = read_csv(here("TCH Data/JS.2.23.23.csv"))

```
# Lactate Turning Point (LTP)
Lactate Turning Point 1 (LTP1) and Lactate Turning Point 2 (LTP2)
the lactate response is divided into three segments. A segmented regression is performed
such that the lactate curve yields two breaking points. The first breaking point, representing
the first rise in lactate above resting levels, is considered as LTP1. The second breaking point,
representing an accelerated lactate accumulation, is then considered as LTP2
(Hofmann & Tschakert, 2017; Hofmann et al., 1997; Pokan et al., 1997).
Caution: this method might require a double-check via a visual inspection, depending in some cases.

```{r echo=FALSE}
results_ltp <- lactate_threshold(
  .data = Data,
  intensity_column = "intensity", 
  lactate_column = "lactate", 
  heart_rate_column = "heart_rate",
  method = "LTP",
  fit = "3rd degree polynomial", 
  include_baseline = TRUE, 
  sport = "cycling",
  plot = TRUE
)
```
```{r echo=FALSE}
LTP1plot = results_ltp$plot[[1]]
LTP2plot = results_ltp$plot[[2]]

dataInterp = LTP2plot$plot_env$data_interpolated
dataFitted = LTP2plot$plot_env$fitted_values
dataRaw = LTP2plot$plot_env$data

LTP1point = LTP1plot$plot_env$breakpoint_y_value
LTP2point = LTP2plot$plot_env$breakpoint_y_value

LTP1pointX= dataFitted[which.min(abs(LTP1point-dataFitted$value)),]
LTP2pointX= dataFitted[which.min(abs(LTP2point-dataFitted$value)),]

x_labels = LTP1plot$plot_env$breaks_and_labels$labels
x_breaks = LTP1plot$plot_env$breaks_and_labels$breaks

```

```{r echo=FALSE, fig.width=10, fig.height=5}
ggplot() + 
  geom_line(data=dataFitted, aes(x=intensity, y=value) ,color = "red") + 
  geom_point(data=dataRaw, aes(x=intensity, y=lactate), size = 5, alpha = 0.2, colour = "gray")+
  geom_point(aes(x=LTP1pointX$intensity, y=LTP1point), size = 5, alpha = 0.3, colour="blue")+
  geom_point(aes(x=LTP2pointX$intensity, y=LTP2point), size = 5, alpha = 0.3, colour="blue")+
  #scale_x_continuous(name="Power (W)", breaks=seq(100,300,20)) +
  scale_x_continuous(name = "Power (W)", breaks = x_breaks, labels = x_labels) +
  ylab("Blood Lactate (mmol/L)")+
  ggtitle("Lactate Turning Points (LTP1 & LTP2)")+
  theme_light()
```

